===
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
===

## Loading and preprocessing the data 

- unzip the file:
```{r}
if (!file.exists("activity.csv")) { 
     unzip("activity.zip")}
```
- load data:
```{r}
activity <- read.csv("activity.csv", header = TRUE)
```
- quick look at the data:
```{r}
str(activity)
head(activity)
```


## What is mean total number of steps taken per day?  
- aggregate needed data and create a histogram of the total number of steps taken each day:
```{r}
totalSteps <- aggregate(steps ~ date, data = activity, sum, na.rm = TRUE)
hist(totalSteps$steps)
```  

- calculate and report the mean and median total number of steps taken per day:
```{r}
stepsmean <- mean(totalSteps$steps)
stepsmedian <-median(totalSteps$steps)
```

The mean value for steps taken per day is `r stepsmean`.
The median value for steps taken per day is `r stepsmedian`.


## What is the average daily activity pattern?
- make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r}
stepsInterval <- aggregate(steps ~ interval, data = activity, mean, na.rm = TRUE)
plot(steps ~ interval, data = stepsInterval, type = "l")
```

- which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
stepsint <- stepsInterval[which.max(stepsInterval$steps), ]$interval
```
It is the `r stepsint`th interval.


## Imputing missing values
- calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{r}
sum(is.na(activity$steps))
```
- devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.

I chose a strategy for filing in all of the missing values with the mean for that 5-minute interval and made a function "interval2steps" to get the mean steps for particular 5-minute interval
```{r}
interval2steps <- function(interval) {
    stepsInterval[stepsInterval$interval == interval, ]$steps
}
```
- create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r}
activityFilled <- activity
count = 0
for (i in 1:nrow(activityFilled)) {
    if (is.na(activityFilled[i, ]$steps)) {
        activityFilled[i, ]$steps <- interval2steps(activityFilled[i, ]$interval)
        count = count + 1
    }
}
```
- make a histogram of the total number of steps taken each day
```{r}
totalSteps2 <- aggregate(steps ~ date, data = activityFilled, sum)
hist(totalSteps2$steps)
```

- calculate and report the mean and median total number of steps taken per day
```{r}
stepsmean2 <- mean(totalSteps2$steps)
stepsmedian2 <- median(totalSteps2$steps)
```
The mean value for steps taken per day is `r stepsmean2`.
The median value for steps taken per day is `r stepsmedian2`.

- Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?

As we put the mean value for that particular 5-min interval, mean doesn't differ, and median differs only slightly.

## Are there differences in activity patterns between weekdays and weekends?
- create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day
```{r}
activityFilled$day = ifelse(as.POSIXlt(as.Date(activityFilled$date))$wday%%6 == 
    0, "weekend", "weekday")
activityFilled$day = factor(activityFilled$day, levels = c("weekday", "weekend"))
```
- make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). The plot should look something like the following, which was creating using simulated data:
```{r}
stepsInterval2 = aggregate(steps ~ interval + day, activityFilled, mean)
library(lattice)
xyplot(steps ~ interval | factor(day), data = stepsInterval2, aspect = 1/2, 
    type = "l")
```


